---
title: "Chap4"
output: html_notebook
---


## Question 1
### (a) prediction range
```{r}
sprintf("the range will be %f to %f", exp(-0.25), exp(0.25))
```
### (b) Regression line
```{r}
# simulate the weight 
# assuming weight has mean 120 pounds, sd of 15 pounds
log_height <- rnorm(1000, mean=log(6), sd=log(1.1))
# assuming a relationship between height and weight
error <- rnorm(1000, mean=0, sd=0.25)
log_weight <- -3.5 + 2.0 * log_height + error
# plot weight and height
plot(log_height,log_weight)
```

```{r}
fit4a <- lm(log_weight ~ log_height)
summary(fit4a)
```

## Question 2
### (a) Clean data
```{r}
d2 = read.dta("../ARM_Data/earnings/heights.dta")
head(d2)
```
```{r}
# drop row that contain na
d2_c <- na.omit(d2[,c("earn", "sex", "height")])
dim(d2_c)
```

### (b) Fit a model
```{r}
# use z scores for height
d2_c$height.z <- scale(d2_c$height, center=TRUE, scale=TRUE)
fit2b <- lm(earn ~ height.z, data=d2_c)
summary(fit2b)
```

```{r}
# plot the relationship between earning and height
plot(d2_c$height.z, d2_c$earn)
curve(cbind(1, x) %*% coef(fit2b), add=TRUE)
```

### (c) Try different models
Try including all predictors 
```{r}
# transform the sex into an indicator variable
d2_c$sex.ind <- d2_c$sex - 1
fit2c_1 <- lm(earn ~ height.z + sex.ind, data=d2_c)
summary(fit2c_1)
```
```{r}
# try log transform the earn
# only fit the ones with earning greater than zero
d2_c_pos <- d2_c[d2_c$earn>0,]
d2_c_pos$log_earn <- log(d2_c_pos$earn)
fit2c_2 <- lm(log_earn ~ height.z + sex.ind, data=d2_c_pos)
summary(fit2c_2)
```

```{r}
# try including interactions
fit2c_3 <- lm(earn ~ height.z + sex.ind + height.z:sex.ind, data=d2_c)
summary(fit2c_3)
```

## Question 4
### (a) scatter plot
```{r}
# load the data
d4 <- read.dta("../ARM_Data/pollution.dta")
head(d4)
```

```{r}
plot(d4$nox, d4$mort, xlab="nitric oxides", ylab="mortality")
```
There seems a positive correlation between mortality and nitric oxides, although some outliers exist.
Fit the model:
```{r}
fit4a <- lm(mort ~ nox, data=d4)
summary(fit4a)
```
Plot the residuals:
```{r}
resid <- residuals(fit4a)
pred <- predict(fit4a)
plot(pred, resid, xlab="Predicted value", ylab="Residuals")
```

### (b) Try transformation
Log-transform the nitric oxides level:
```{r}
d4_t <- d4
d4_t$log_nox <- log(d4$nox)
fit4b <- lm(mort ~ log_nox, data=d4_t)
summary(fit4b)
```
Log-tranform  mortality:
```{r}
d4_t$log_mort <- log(d4$mort)
fit4b2 <- lm(log_mort ~ nox, data=d4_t)
summary(fit4b2)
```
Log-transform both mortality and nitric acid
```{r}
fit4b3 <- lm(log_mort ~ log_nox, data=d4_t)
summary(fit4b3)
```
Plot the log_mort vs log_nox:
```{r}
plot(d4_t$log_nox, d4_t$log_mort)
curve(cbind(1,x) %*% coef(fit4b3), add=TRUE)
```

### (c) Intepretation
After log-transformation of both nitrix acid and mortality, coefficient means how much percent mortality will increase when the nitric acid increase 1%, here it's exp(0.0159), which should be very close to 0.0159%. 
If just log-transform the mortality, coefficient means how many percents mortality will increase when increase nitric acid by 1 unit. 

### (d) Multi-predictor model
```{r}
d4_t$log_so2 <- log(d4_t$so2)
d4_t$log_hc <- log(d4_t$hc)
fit4d <- lm(log_mort ~ log_nox + log_so2 + log_hc, data=d4_t)
summary(fit4d)
```
```{r}
par(mfrow=c(1,3))
plot(d4_t$log_nox, d4_t$log_mort, xlab="log(NOX)", ylab="log(mortality)")
curve(cbind(1,x,mean(d4_t$log_so2),mean(d4_t$log_hc)) %*% coef(fit4d), add=TRUE)
plot(d4_t$log_so2, d4_t$log_mort, xlab="log(SO2)", ylab="log(mortality)")
curve(cbind(1,mean(d4_t$log_nox),x,mean(d4_t$log_hc)) %*% coef(fit4d), add=TRUE)
plot(d4_t$log_nox, d4_t$log_mort, xlab="log(HC)", ylab="log(mortality)")
curve(cbind(1,mean(d4_t$log_nox),mean(d4_t$log_hc),x) %*% coef(fit4d), add=TRUE)
```
### (e) cross-validataion
Fit half of the data:
```{r}
d4_size <- dim(d4_t)[1]
d4_t_half1 <- d4_t[1:(d4_size/2),]
d4_t_half2 <- d4_t[(d4_size/2):d4_size,]
fit4e <- lm(log_mort ~ log_nox + log_so2 + log_hc, data=d4_t_half1)
summary(fit4e)
```
Make predictions: 
```{r}
d4_pred <- predict(fit4e, d4_t_half2)
plot(d4_pred, d4_t_half2$log_mort, xlab="log(Predicted mortality)", ylab="log(Actual mortality)", xlim=c(6.6,7.1), ylim=c(6.6, 7.1))
abline(0,1)
```


